---
title: "PI and HRV"
author: "Celine"
date: "August 7, 2017"
output: word_document
---
## PI behavioral results

```{r, echo=FALSE,warning=FALSE, include=FALSE}
options(contrasts=c("contr.sum", "contr.poly"))
library(foreign)
library(psych)
library(reshape2)
library(ggplot2)
library(car)
library(ez)
library(Rmisc)
library(gmodels)
library(afex)
library(pander)
library(wesanderson)

# for AB acquisition
sessList <- list(136,137,138,139)#list session nubmers here
ratList <- list(101,102,103,104,105,106,107,108,109,110,112)#list rat number here
x <- length(sessList)
y <- length(ratList)
rowsNeeded <- (x*y)

ABsummary <- matrix(data = NA, nrow = rowsNeeded, ncol=3)
colnames(ABsummary)<- c('ratID', 'session', 'accuracy')
i = 1
for (sessNum in sessList){
  for (rat in ratList) {
filename <- paste("U3_",rat,"_",sessNum,".csv", sep="")
    print(filename)
    currentData <- read.csv(filename, header=TRUE, sep="\t")
    currentData$resp2 <- ifelse(currentData$Response == "correct", 1,0)
    currentAB <- subset(currentData, trialType == "AB")
    ABacc <- mean(currentAB$resp2)
    ABsummary[i,0:3] <- c(rat, sessNum, ABacc)
    i=i+1
    
  }
}
ABsummary <- data.frame(ABsummary)
ABsummary$ratID <- as.factor(ABsummary$ratID)
ABsummary$session <- as.factor(ABsummary$session)
ABsummaryTable <- aggregate(accuracy~session, data=ABsummary, FUN=(mean))
tempTable <- aggregate(accuracy~session, data=ABsummary, FUN=(sd))
ABsummaryTable<- as.data.frame(ABsummaryTable)
ABsummaryTable$sd <- tempTable$accuracy*100
tempTable$accuracy <- tempTable$accuracy/sqrt(y)
ABsummaryTable$se <- tempTable$accuracy*100
ABsummaryTable$accuracy <- ABsummaryTable$accuracy * 100

#for full task
sessList <- list(140,141,142,143,144)#list session nubmers here
ratList <- list(101,102,103,104,105,106,107,108,109,110,112)#list rat number here
x <- length(sessList)
y <- length(ratList)
rowsNeeded <- (x*y)
FULLsummary <- matrix(data = NA, nrow = rowsNeeded, ncol=5)
colnames(FULLsummary)<- c('ratID', 'session', 'AB_acc', 'CA_acc', 'DE_acc')

i = 1
for (sessNum in sessList){
  for (rat in ratList) {
    filename <- paste("U3_",rat,"_",sessNum,".csv", sep="")
    print(filename)
    currentData <- read.csv(filename, header=TRUE, sep="\t")
    currentData$resp2 <- ifelse(currentData$Response == "correct", 1,0)
    currentAB <- subset(currentData, trialType == "AB")
    ABacc <- mean(currentAB$resp2)
    currentDE <- subset(currentData, trialType == "DE")
    DEacc <- mean(currentDE$resp2)
    currentAC <- subset(currentData, trialType == "CA")
    ACacc <- mean(currentAC$resp2)
    FULLsummary[i,0:5] <- c(rat, sessNum, ABacc, ACacc, DEacc)
    i=i+1
    
  }
}
#prepare wide and long form data frames for full task
FULLfinalSummaryW <- data.frame(FULLsummary)
#FULLfinalSummaryW$PI_effect <- FULLfinalSummaryW$DE_acc - FULLfinalSummaryW$CA_acc
z = rowsNeeded*3
FULLfinalSummaryL <- reshape(FULLfinalSummaryW, varying = c('AB_acc', 'CA_acc', 'DE_acc'), v.names="accuracy",
                         timevar='trialType',times=c('AB', 'CA', 'DE'), new.row.names=1:z, direction = 'long')
FULLfinalSummaryL$ratID <-as.factor(FULLfinalSummaryL$ratID)
FULLfinalSummaryL$session <-as.factor(FULLfinalSummaryL$session)
FULLfinalSummaryL$trialType <- as.factor(FULLfinalSummaryL$trialType)
FULLfinalSummaryW$ratID <-as.factor(FULLfinalSummaryW$ratID)
FULLfinalSummaryW$session <-as.factor(FULLfinalSummaryW$session)

#making summary table for full task
FULLsummaryTable <- aggregate(accuracy~trialType, data=FULLfinalSummaryL, FUN=(mean))
tempTable <- aggregate(accuracy~trialType, data=FULLfinalSummaryL, FUN=(sd))
FULLsummaryTable<- as.data.frame(FULLsummaryTable)
FULLsummaryTable$sd <- tempTable$accuracy
tempTable$accuracy <- tempTable$accuracy/sqrt(y)
FULLsummaryTable$se <- tempTable$accuracy
FULLsummaryTable$accuracy <- FULLsummaryTable$accuracy

```
### Descriptives and plots
```{r, echo=FALSE}
#part 1 - AB acquisition
#check distribution
pander(describe(ABsummary))
hist(ABsummary$accuracy,main = ('Distribution of Accuracy during AB acquisiton, \nall sessions'),xlab = ('AB accuracy'))
#distribution is actually a bit strange, though all above chance - also pretty small number of rats
plot(accuracy~session, data=ABsummary, main=('AB accuracy by session, AB acquisition'))
#part 2 - full task
pander(describe(FULLfinalSummaryW))
hist(FULLfinalSummaryL$accuracy, main = ('Distribution of Overall Accuracy, \nFull Task Days'),xlab=('Overall Accuracy'))
#accuracy for the full task is normally distributed 
plot(accuracy~session, data=FULLfinalSummaryL,main=('Overall accuracy by session, \nfull task days'))
#looks like session didn't ahve much effect
p1= ggplot(data=FULLfinalSummaryL, aes(trialType,accuracy))
p1+geom_boxplot(aes(fill=trialType)) +
  coord_cartesian(ylim=c(0,1))+
  ggtitle("Accuracy by Trial type, \naveraged across full task days") +
  labs(x="Trial Type",y="Accuracy")+
  scale_fill_manual(values=wes_palette("FantasticFox"))+
  guides(fill=guide_legend(title="trial type"))

ggplot(FULLsummaryTable, aes(x=trialType, y=accuracy, fill=trialType)) + 
  geom_bar(stat="identity", color="black", 
           position=position_dodge()) +
  coord_cartesian(ylim=c(0,1))+
  geom_errorbar(aes(ymin=accuracy-se, ymax=accuracy+se), width=.2,
                position=position_dodge(.9)) +
  scale_fill_manual(values=wes_palette("FantasticFox"))+
  ggtitle("Accuracy by trial trype, \naveraged across days")+
  guides(fill=guide_legend(title="trial type")) +
  labs(x="Trial Type",y="Accuracy")

FULLsummaryTable <- aggregate(accuracy~trialType+session, data=FULLfinalSummaryL, FUN=(mean))
tempTable <- aggregate(accuracy~trialType+session, data=FULLfinalSummaryL, FUN=(sd))
FULLsummaryTable<- as.data.frame(FULLsummaryTable)
FULLsummaryTable$sd <- tempTable$accuracy
tempTable$accuracy <- tempTable$accuracy/sqrt(y)
FULLsummaryTable$se <- tempTable$accuracy
FULLsummaryTable$accuracy <- FULLsummaryTable$accuracy

subSummaryTable <-subset(FULLsummaryTable, trialType != 'AB')

ggplot(FULLsummaryTable, aes(x=session,y=accuracy,colour = trialType))+
  geom_line(aes(group=trialType))+
  geom_point()+geom_errorbar(aes(ymin=accuracy-se, ymax=accuracy+se),width=.1,position=position_dodge(.05))+
  coord_cartesian(ylim=c(0,1))+
  guides(colour=guide_legend(title="trial type"))+
  scale_color_manual(values=wes_palette("FantasticFox"))+
  ggtitle('Accuracy by Trial Type and Session, \nFull Task Days')
```

### Inferential statistics
Within-subjects ANOVA for trial type by session for full task days
```{r, echo=FALSE}
aov2 <- aov_ez("ratID","accuracy",data=FULLfinalSummaryL,within = c("trialType","session"))
pander(summary(aov2))
```

Post hoc pairwise t-tests for trial type with Holm correction for multiple testing
```{r, echo=FALSE}
lsmeans(aov2, "trialType", contr = "pairwise", adjust = "holm")
```

Fine-slice simple main effect of trial type within each session separately 
```{r, echo=FALSE}
sess140 <- subset(FULLfinalSummaryL, session == "140")
sess141 <- subset(FULLfinalSummaryL, session == "141")
sess142 <- subset(FULLfinalSummaryL, session == "142")
sess143 <- subset(FULLfinalSummaryL, session == "143")
sess144 <- subset(FULLfinalSummaryL, session == "144")
aov4 <- aov_ez("ratID","accuracy",data=sess140,within=c("trialType"))
aov5 <- aov_ez("ratID","accuracy",data=sess141,within=c("trialType"))
aov6 <- aov_ez("ratID","accuracy",data=sess142,within=c("trialType"))
aov7 <- aov_ez("ratID","accuracy",data=sess143,within=c("trialType"))
aov8 <- aov_ez("ratID","accuracy",data=sess144,within=c("trialType"))
pander(summary(aov4))
pander(summary(aov5))
pander(summary(aov6))
pander(summary(aov7))
pander(summary(aov8))
```

Pairwise t-tests of trial type within each session, with Holm correction for multiple testing
```{r, echo=FALSE}
pander(summary(lsmeans(aov4, "trialType", contr = "pairwise", adjust = "holm")))
pander(summary(lsmeans(aov5, "trialType", contr = "pairwise", adjust = "holm")))
pander(summary(lsmeans(aov6, "trialType", contr = "pairwise", adjust = "holm")))
pander(summary(lsmeans(aov7, "trialType", contr = "pairwise", adjust = "holm")))
pander(summary(lsmeans(aov8, "trialType", contr = "pairwise", adjust = "holm")))
```

##HRV results

### Descriptives

```{r, echo=FALSE}
hrv_data <- read.csv("HRV_PI.csv")

hr_by_rat <- dcast(hrv_data, rat~session,value.var="avg_HR")
hr_by_rat <- data.frame(hr_by_rat)
colnames(hr_by_rat)<- c("ratID","day_1","day_2")


rmssd_by_rat <- dcast(hrv_data, rat~session,value.var="rmssd")
rmssd_by_rat <- data.frame(rmssd_by_rat)
colnames(rmssd_by_rat)<-c("ratID","day_1","day_2")
```
Descriptive statistics for heart rate on each day
```{r, echo=FALSE}
pander(describe(hr_by_rat))
```
Descriptive statistics for heart rate on each day
```` {r, echo=FALSE}
pander(describe(rmssd_by_rat))
```

###Comparing day 1 and day 2 - RMSSD
Regression results for relationship between RMSSD on day 1 and RMSSD on day 2 for each rat
```{r, echo=FALSE}
lm1 <- lm(day_1~day_2,data=rmssd_by_rat)
```
R squared: 
```{r,echo=FALSE}
summary(lm1)$r.squared
plot(day_1~day_2,data=rmssd_by_rat, main="RMSSD for each rat, day 1 vs. day 2", xlab="RMSSD day 2",ylab="RMSSD day 1")
abline(lm(day_1~day_2,data=rmssd_by_rat))
```
Regression results for relationship between heart rate on day 1 and heart rate on day 2 for each rat
```{r, echo=FALSE}
lm2 <- lm(day_1~day_2,data=hr_by_rat)
```
R squared:
```{r, echo=FALSE}
summary(lm2)$r.squared
plot(day_1~day_2,data=hr_by_rat, main="heart rate for each rat, day 1 vs. day 2", xlab="HR day 2",ylab="HR day 1")
abline(lm(day_1~day_2,data=hr_by_rat))
```

## Relationship between PI and HRV
```{r, echo=FALSE}
#making a master data frame
meltedDF <- melt(FULLfinalSummaryW)
masterDataFrame <- dcast(meltedDF, ratID~variable, value.var = "value", fun.aggregate = mean)
masterDataFrame <- as.data.frame(masterDataFrame)
masterDataFrame$HR_day_1 <- hr_by_rat$day_1
masterDataFrame$HR_day_2 <- hr_by_rat$day_2
masterDataFrame$rmssd_day_1 <- rmssd_by_rat$day_1
masterDataFrame$rmssd_day_2 <- rmssd_by_rat$day_2
deltaValue <- matrix(data=NA,nrow=length(ratList),ncol = 5)
colnames(deltaValue)<-c("ratID","DE_140","DE_141","CA_140","CA_141")
i=1
for(rat in ratList){
subset <- subset(FULLfinalSummaryW, ratID == rat)
DE_140 <- subset[1,5]
DE_141  <- subset[2,5]
CA_140 <- subset[1,4]
CA_141 <- subset[2,4]
deltaValue[i,]<-c(rat,DE_140,DE_141,CA_140,CA_141)
i=i+1}
masterDataFrame <- merge(masterDataFrame,deltaValue,by="ratID")
masterDataFrame$meanRMSSD<-rowMeans(masterDataFrame[7:8])
masterDataFrame$meanHR<-rowMeans(masterDataFrame[5:6])
masterDataFrame$deltaDE <- masterDataFrame$DE_141 - masterDataFrame$DE_140
masterDataFrame$deltaCA <- masterDataFrame$CA_141 - masterDataFrame$CA_140
masterDataFrame$diff_delta <- masterDataFrame$deltaDE - masterDataFrame$deltaCA
masterDataFrame$PI_effect <- masterDataFrame$DE_acc - masterDataFrame$CA_acc
```

Multivariate multiple regression examining the reltaionship between
* Predictors
  + Mean HR
  + Mean RMSSD
  + RMSSD day 2 (data seemed to be cleaner on day 2)
* Outcomes
  + AB accuracy for all full task days
  + DE accuracy for all full task days
  + CA accuracy for all full task days
  + "PI effect," i.e., DE accuracy - CA accuracy for all full task days
  + DE accuracy on full task day 2
  + CA accuracy on full task day 2
  + Change in accuracy from full task day 1 to full task day 2 in DE vs. CA

```{r, echo=FALSE}
lm1 <- lm(cbind(CA_acc,DE_acc,PI_effect,DE_141,CA_141,diff_delta)~meanHR+meanRMSSD+rmssd_day_2, data=masterDataFrame)
pander(summary(lm1))
```
  


#Replication of PI and PI/HRV analyses above with only rats who successfully learned the AB pair.
Succesful learning defined as reaching an average accuracy of 70% or more in the last two days of AB acquisition.
```{r, echo=FALSE, warning=FALSE, include=FALSE}
AB_wide <- dcast(ABsummary, ratID~session, value.var = "accuracy")
AB_wide$last_two_days <- (AB_wide$`138`+ AB_wide$`139`)/2
keepRats <- subset(AB_wide, last_two_days >=.70)
keepRats <- array(keepRats$ratID)
keepRats

# for AB acquisition
sessList <- list(136,137,138,139)#list session nubmers here
ratList <- list(101,102,107,108,109,112)#list rat number here
x <- length(sessList)
y <- length(ratList)
rowsNeeded <- (x*y)

ABsummary <- matrix(data = NA, nrow = rowsNeeded, ncol=3)
colnames(ABsummary)<- c('ratID', 'session', 'accuracy')
i = 1
for (sessNum in sessList){
  for (rat in ratList) {
filename <- paste("U3_",rat,"_",sessNum,".csv", sep="")
    print(filename)
    currentData <- read.csv(filename, header=TRUE, sep="\t")
    currentData$resp2 <- ifelse(currentData$Response == "correct", 1,0)
    currentAB <- subset(currentData, trialType == "AB")
    ABacc <- mean(currentAB$resp2)
    ABsummary[i,0:3] <- c(rat, sessNum, ABacc)
    i=i+1
    
  }
}
ABsummary <- data.frame(ABsummary)
ABsummary$ratID <- as.factor(ABsummary$ratID)
ABsummary$session <- as.factor(ABsummary$session)
ABsummaryTable <- aggregate(accuracy~session, data=ABsummary, FUN=(mean))
tempTable <- aggregate(accuracy~session, data=ABsummary, FUN=(sd))
ABsummaryTable<- as.data.frame(ABsummaryTable)
ABsummaryTable$sd <- tempTable$accuracy*100
tempTable$accuracy <- tempTable$accuracy/sqrt(y)
ABsummaryTable$se <- tempTable$accuracy*100
ABsummaryTable$accuracy <- ABsummaryTable$accuracy * 100

#for full task
sessList <- list(140,141,142,143,144)#list session nubmers here
ratList <- list(101,102,107,108,109,112)
x <- length(sessList)
y <- length(ratList)
rowsNeeded <- (x*y)
FULLsummary <- matrix(data = NA, nrow = rowsNeeded, ncol=5)
colnames(FULLsummary)<- c('ratID', 'session', 'AB_acc', 'CA_acc', 'DE_acc')

i = 1
for (sessNum in sessList){
  for (rat in ratList) {
    filename <- paste("U3_",rat,"_",sessNum,".csv", sep="")
    print(filename)
    currentData <- read.csv(filename, header=TRUE, sep="\t")
    currentData$resp2 <- ifelse(currentData$Response == "correct", 1,0)
    currentAB <- subset(currentData, trialType == "AB")
    ABacc <- mean(currentAB$resp2)
    currentDE <- subset(currentData, trialType == "DE")
    DEacc <- mean(currentDE$resp2)
    currentAC <- subset(currentData, trialType == "CA")
    ACacc <- mean(currentAC$resp2)
    FULLsummary[i,0:5] <- c(rat, sessNum, ABacc, ACacc, DEacc)
    i=i+1
    
  }
}
#prepare wide and long form data frames for full task
FULLfinalSummaryW <- data.frame(FULLsummary)
#FULLfinalSummaryW$PI_effect <- FULLfinalSummaryW$DE_acc - FULLfinalSummaryW$CA_acc
z = rowsNeeded*3
FULLfinalSummaryL <- reshape(FULLfinalSummaryW, varying = c('AB_acc', 'CA_acc', 'DE_acc'), v.names="accuracy",
                         timevar='trialType',times=c('AB', 'CA', 'DE'), new.row.names=1:z, direction = 'long')
FULLfinalSummaryL$ratID <-as.factor(FULLfinalSummaryL$ratID)
FULLfinalSummaryL$session <-as.factor(FULLfinalSummaryL$session)
FULLfinalSummaryL$trialType <- as.factor(FULLfinalSummaryL$trialType)
FULLfinalSummaryW$ratID <-as.factor(FULLfinalSummaryW$ratID)
FULLfinalSummaryW$session <-as.factor(FULLfinalSummaryW$session)

#making summary table for full task
FULLsummaryTable <- aggregate(accuracy~trialType, data=FULLfinalSummaryL, FUN=(mean))
tempTable <- aggregate(accuracy~trialType, data=FULLfinalSummaryL, FUN=(sd))
FULLsummaryTable<- as.data.frame(FULLsummaryTable)
FULLsummaryTable$sd <- tempTable$accuracy
tempTable$accuracy <- tempTable$accuracy/sqrt(y)
FULLsummaryTable$se <- tempTable$accuracy
FULLsummaryTable$accuracy <- FULLsummaryTable$accuracy

```
### Descriptives and plots
```{r, echo=FALSE}
#part 1 - AB acquisition
#check distribution
pander(describe(ABsummary))
hist(ABsummary$accuracy,main = ('high acheivers Distribution of Accuracy \nduring AB acquisiton, all sessions'),xlab = ('AB accuracy'))
#distribution is actually a bit strange, though all above chance - also pretty small number of rats
plot(accuracy~session, data=ABsummary, main=('high acheivers AB accuracy by session, \nAB acquisition'))
#part 2 - full task
pander(describe(FULLfinalSummaryW))
hist(FULLfinalSummaryL$accuracy, main = ('high acheivers Distribution of Overall Accuracy, \nfull fask days'),xlab=('Overall Accuracy'))
#accuracy for the full task is normally distributed 
plot(accuracy~session, data=FULLfinalSummaryL,main=('high acheivers Overall accuracy by session, \nfull task days'))
#looks like session didn't ahve much effect
p1= ggplot(data=FULLfinalSummaryL, aes(trialType,accuracy))
p1+geom_boxplot(aes(fill=trialType)) +
  coord_cartesian(ylim=c(0,1))+
  ggtitle("high acheivers Accuracy by Trial type, \naveraged across full task days") +
  labs(x="Trial Type",y="Accuracy")+
  scale_fill_manual(values=wes_palette("FantasticFox"))+
  guides(fill=guide_legend(title="trial type"))

ggplot(FULLsummaryTable, aes(x=trialType, y=accuracy, fill=trialType)) + 
  geom_bar(stat="identity", color="black", 
           position=position_dodge()) +
  coord_cartesian(ylim=c(0,1))+
  geom_errorbar(aes(ymin=accuracy-se, ymax=accuracy+se), width=.2,
                position=position_dodge(.9)) +
  scale_fill_manual(values=wes_palette("FantasticFox"))+
  ggtitle("high acheivers Accuracy by trial trype, \naveraged across days")+
  guides(fill=guide_legend(title="trial type")) +
  labs(x="Trial Type",y="Accuracy")

FULLsummaryTable <- aggregate(accuracy~trialType+session, data=FULLfinalSummaryL, FUN=(mean))
tempTable <- aggregate(accuracy~trialType+session, data=FULLfinalSummaryL, FUN=(sd))
FULLsummaryTable<- as.data.frame(FULLsummaryTable)
FULLsummaryTable$sd <- tempTable$accuracy
tempTable$accuracy <- tempTable$accuracy/sqrt(y)
FULLsummaryTable$se <- tempTable$accuracy
FULLsummaryTable$accuracy <- FULLsummaryTable$accuracy

subSummaryTable <-subset(FULLsummaryTable, trialType != 'AB')

ggplot(FULLsummaryTable, aes(x=session,y=accuracy,colour = trialType))+
  geom_line(aes(group=trialType))+
  geom_point()+geom_errorbar(aes(ymin=accuracy-se, ymax=accuracy+se),width=.1,position=position_dodge(.05))+
  coord_cartesian(ylim=c(0,1))+
  guides(colour=guide_legend(title="trial type"))+
  scale_color_manual(values=wes_palette("FantasticFox"))+
  ggtitle('high acheivers Accuracy by Trial Type and Session, \nfull task days')
```

### Inferential statistics
Within-subjects ANOVA for trial type by session for full task days
```{r, echo=FALSE}
aov2 <- aov_ez("ratID","accuracy",data=FULLfinalSummaryL,within = c("trialType","session"))
pander(summary(aov2))
```

Post hoc pairwise t-tests for trial type with Holm correction for multiple testing
```{r, echo=FALSE}
lsmeans(aov2, "trialType", contr = "pairwise", adjust = "holm")
```

Fine-slice simple main effect of trial type within each session separately 
```{r, echo=FALSE}
sess140 <- subset(FULLfinalSummaryL, session == "140")
sess141 <- subset(FULLfinalSummaryL, session == "141")
sess142 <- subset(FULLfinalSummaryL, session == "142")
sess143 <- subset(FULLfinalSummaryL, session == "143")
sess144 <- subset(FULLfinalSummaryL, session == "144")
aov4 <- aov_ez("ratID","accuracy",data=sess140,within=c("trialType"))
aov5 <- aov_ez("ratID","accuracy",data=sess141,within=c("trialType"))
aov6 <- aov_ez("ratID","accuracy",data=sess142,within=c("trialType"))
aov7 <- aov_ez("ratID","accuracy",data=sess143,within=c("trialType"))
aov8 <- aov_ez("ratID","accuracy",data=sess144,within=c("trialType"))
pander(summary(aov4))
pander(summary(aov5))
pander(summary(aov6))
pander(summary(aov7))
pander(summary(aov8))
```

Pairwise t-tests of trial type within each session, with Holm correction for multiple testing
```{r, echo=FALSE}
pander(summary(lsmeans(aov4, "trialType", contr = "pairwise", adjust = "holm")))
pander(summary(lsmeans(aov5, "trialType", contr = "pairwise", adjust = "holm")))
pander(summary(lsmeans(aov6, "trialType", contr = "pairwise", adjust = "holm")))
pander(summary(lsmeans(aov7, "trialType", contr = "pairwise", adjust = "holm")))
pander(summary(lsmeans(aov8, "trialType", contr = "pairwise", adjust = "holm")))