---
title: "PI and HRV"
author: "Celine"
date: "August 7, 2017"
output: word_document
---
## PI behavioral results


```{r, echo=FALSE}
options(contrasts=c("contr.sum", "contr.poly"))
library(foreign)
library(psych)
library(reshape2)
library(ggplot2)
library(car)
library(ez)
library(Rmisc)
library(gmodels)
library(package)
library(afex)
library(pander)

# for AB acquisition
sessList <- list(136,137,138,139)#list session nubmers here
ratList <- list(101,102,103,104,105,106,107,108,109,110,112)#list rat number here
x <- length(sessList)
y <- length(ratList)
rowsNeeded <- (x*y)

ABsummary <- matrix(data = NA, nrow = rowsNeeded, ncol=3)
colnames(ABsummary)<- c('ratID', 'session', 'accuracy')
i = 1
for (sessNum in sessList){
  for (rat in ratList) {
filename <- paste("U3_",rat,"_",sessNum,".csv", sep="")
    print(filename)
    currentData <- read.csv(filename, header=TRUE, sep="\t")
    currentData$resp2 <- ifelse(currentData$Response == "correct", 1,0)
    currentAB <- subset(currentData, trialType == "AB")
    ABacc <- mean(currentAB$resp2)
    ABsummary[i,0:3] <- c(rat, sessNum, ABacc)
    i=i+1
    
  }
}
ABsummary <- data.frame(ABsummary)
ABsummary$ratID <- as.factor(ABsummary$ratID)
ABsummary$session <- as.factor(ABsummary$session)
ABsummaryTable <- aggregate(accuracy~session, data=ABsummary, FUN=(mean))
tempTable <- aggregate(accuracy~session, data=ABsummary, FUN=(sd))
ABsummaryTable<- as.data.frame(ABsummaryTable)
ABsummaryTable$sd <- tempTable$accuracy*100
tempTable$accuracy <- tempTable$accuracy/sqrt(y)
ABsummaryTable$se <- tempTable$accuracy*100
ABsummaryTable$accuracy <- ABsummaryTable$accuracy * 100

#for full task
sessList <- list(140,141,142,143,144)#list session nubmers here
ratList <- list(101,102,103,104,105,106,107,108,109,110,112)#list rat number here
x <- length(sessList)
y <- length(ratList)
rowsNeeded <- (x*y)
FULLsummary <- matrix(data = NA, nrow = rowsNeeded, ncol=5)
colnames(FULLsummary)<- c('ratID', 'session', 'AB_acc', 'AC_acc', 'DE_acc')

i = 1
for (sessNum in sessList){
  for (rat in ratList) {
    filename <- paste("U3_",rat,"_",sessNum,".csv", sep="")
    print(filename)
    currentData <- read.csv(filename, header=TRUE, sep="\t")
    currentData$resp2 <- ifelse(currentData$Response == "correct", 1,0)
    currentAB <- subset(currentData, trialType == "AB")
    ABacc <- mean(currentAB$resp2)
    currentDE <- subset(currentData, trialType == "DE")
    DEacc <- mean(currentDE$resp2)
    currentAC <- subset(currentData, trialType == "CA")
    ACacc <- mean(currentAC$resp2)
    FULLsummary[i,0:5] <- c(rat, sessNum, ABacc, ACacc, DEacc)
    i=i+1
    
  }
}
#prepare wide and long form data frames for full task
FULLfinalSummaryW <- data.frame(FULLsummary)
#FULLfinalSummaryW$PI_effect <- FULLfinalSummaryW$DE_acc - FULLfinalSummaryW$AC_acc
z = rowsNeeded*3
FULLfinalSummaryL <- reshape(FULLfinalSummaryW, varying = c('AB_acc', 'AC_acc', 'DE_acc'), v.names="accuracy",
                         timevar='trialType',times=c('AB', 'CA', 'DE'), new.row.names=1:z, direction = 'long')
FULLfinalSummaryL$ratID <-as.factor(FULLfinalSummaryL$ratID)
FULLfinalSummaryL$session <-as.factor(FULLfinalSummaryL$session)
FULLfinalSummaryL$trialType <- as.factor(FULLfinalSummaryL$trialType)
FULLfinalSummaryW$ratID <-as.factor(FULLfinalSummaryW$ratID)
FULLfinalSummaryW$session <-as.factor(FULLfinalSummaryW$session)

#making summary table for full task
FULLsummaryTable <- aggregate(accuracy~trialType, data=FULLfinalSummaryL, FUN=(mean))
tempTable <- aggregate(accuracy~trialType, data=FULLfinalSummaryL, FUN=(sd))
FULLsummaryTable<- as.data.frame(FULLsummaryTable)
FULLsummaryTable$sd <- tempTable$accuracy
tempTable$accuracy <- tempTable$accuracy/sqrt(y)
FULLsummaryTable$se <- tempTable$accuracy
FULLsummaryTable$accuracy <- FULLsummaryTable$accuracy

```

``` {r, echo=F}

```

HRV results
### Descriptives

```{r, echo=FALSE}
hrv_data <- read.csv("HRV_PI.csv")

hr_by_rat <- dcast(hrv_data, rat~session,value.var="avg_HR")
hr_by_rat <- data.frame(hr_by_rat)
colnames(hr_by_rat)<- c("ratID","day_1","day_2")


rmssd_by_rat <- dcast(hrv_data, rat~session,value.var="rmssd")
rmssd_by_rat <- data.frame(rmssd_by_rat)
colnames(rmssd_by_rat)<-c("ratID","day_1","day_2")
```
Descriptive statistics for heart rate on each day
```{r, echo=FALSE}
pander(describe(hr_by_rat))
```
Descriptive statistics for heart rate on each day
```` {r, echo=FALSE}
pander(describe(rmssd_by_rat))
```
###Comparing day 1 and day 2 - RMSSD

Regression results for relationship between RMSSD on day 1 and RMSSD on day 2 for each rat
```{r, echo=FALSE}
lm1 <- lm(day_1~day_2,data=rmssd_by_rat)
```
R squared: 
```{r,echo=FALSE}
summary(lm1)$r.squared
plot(day_1~day_2,data=rmssd_by_rat, main="RMSSD for each rat, day 1 vs. day 2", xlab="RMSSD day 2",ylab="RMSSD day 1")
abline(lm(day_1~day_2,data=rmssd_by_rat))
```

Regression results for relationship between heart rate on day 1 and heart rate on day 2 for each rat
```{r, echo=FALSE}
lm2 <- lm(day_1~day_2,data=hr_by_rat)
```
R squared:
```{r, echo=FALSE}
summary(lm2)$r.squared
plot(day_1~day_2,data=hr_by_rat, main="heart rate for each rat, day 1 vs. day 2", xlab="HR day 2",ylab="HR day 1")
abline(lm(day_1~day_2,data=hr_by_rat))
```
